---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(sqlitemeta)
```

# sqlitemeta

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->

`{sqlitemeta}` is a tool for parsing SQLite `CREATE TABLE` statements.

[SQLite](https://sqlite.org) stores and updates information about the columns
it contains in a stored SQL statement as part of the database.

This stored statement is an SQL `CREATE TABLE` operation describing how 
you could the create the table in question.

So if you want to learn about the structure of an existing database, 
parsing this `CREATE TABLE` statement is a key part of understanding 
how it was constructed.

This package is wrapper around the C libary [sqlite-createtable-parser](https://github.com/marcobambini/sqlite-createtable-parser).


## What's in the box

* `parse_sql()` will parse a SQLite `CREATE TABLE` statement (provided as a single
   character string) and return a named list of information.


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/sqlitemeta) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/sqlitemeta')
```



## Example 1 - Parsing a provided `CREATE TABLE` statement


```{r example}
library(sqlitemeta)

sql <- r"(
CREATE TABLE contact_groups(                
   contact_id INTEGER PRIMARY KEY,      
   group_id INTEGER DEFAULT 999,
   details VARCHAR(20),
   PRIMARY KEY (contact_id, group_id),      
   FOREIGN KEY (contact_id)                 
      REFERENCES contacts (contact_id)      
         ON DELETE CASCADE                  
         ON UPDATE NO ACTION,               
   FOREIGN KEY (group_id)                   
      REFERENCES groups (group_id)          
         ON DELETE CASCADE                  
         ON UPDATE NO ACTION                
) STRICT;
)"

parse_sql(sql)
```



## Example 2 - Extracting `CREATE TABLE` statement from existing SQLite database


```{r}
library(sqlitemeta)
library(RSQLite)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create an in-memory SQLite database and add 'mtcars' data as a table
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
con <- dbConnect(RSQLite::SQLite(), ":memory:")
dbWriteTable(con, "mtcars", mtcars, overwrite = TRUE)
dbListTables(con)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Query Sqlite for the 'CREATE TABLE' statement used to 
# create the 'mtcars' table in the database.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
qry <- "SELECT sql FROM sqlite_master WHERE name = 'mtcars';"
res <- dbSendQuery(con, qry)
res <- dbFetch(res)
create_table_sql <- res$sql[[1]] 

cat(create_table_sql, sep = "\n")

parse_sql(create_table_sql)
```



## Related Software

* [RSQLite](https://cran.dev/RSQLite)

## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
